---
title: "Exercise_5"
author: "Eleanor Lindsey"
format: html
editor: visual
---

Attach the `palmerspenguins` package

1\. Examine at the dataset using the ?Help page

2\. what is the class of the penguins dataset?

3\. what is the structure of the penguins dataset?

4\. what are the dimensions of the penguins dataset?

5\. what are the column names of the penguins dataset?

6\. what type of data is `flipper_length_mm` and `Island`?

7\. what is the mean flipper length of the penguins?

8\. what is the standard deviation of flipper length in the penguins?

9\. what is the median body mass of the penguins?

10\. what is the Island of the 100th penguin?

```{r}
library('readr')
read_csv('~/Documents/ESS330-Quantitative-Reasoning/github/exercise-5-csu/penguins.csv')
```

```{r}
library(palmerpenguins)

class(penguins)#what is the class of the penguins dataset?
structure(penguins)#what is the structure of the penguins dataset?
dim(penguins)#what are the dimensions of the penguins dataset? 
class(penguins$flipper_length_mm)#what type of data is `flipper_length_mm` and `Island`?
class(penguins$island)#what type of data is `flipper_length_mm` and `Island`?
mean(penguins$flipper_length_mm, na.rm = TRUE)#what is the mean flipper length of the penguins?
sd(penguins$flipper_length_mm, na.rm = TRUE)#what is the standard deviation of flipper length in the penguins?
median(penguins$body_mass_g, na.rm=TRUE)#what is the median body mass of the penguins?
penguins$island[100]#what is the Island of the 100th penguin?
#Column names for the penguins data set
#"species", "island", "bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g", "sex", "year"
```
