---
title: "Exercise-11-csu"
format: 
  html:
    self-contained: true
editor: visual
---

## In this assignment, you will again analyze the `airquality` dataset using various statistical tests, data transformation techniques, and regression modeling. Follow the step-by-step guiding questions to complete the analysis in a qmd.

## **Part 1: Normality Testing**

1.  Load the airquality dataset in R. What does this dataset represent? Explore its structure using functions like `str()` and `summary()`.

    ```{r}
    airquality <-airquality
    str(airquality)
    summary(airquality)
    ```

2.  Perform a Shapiro-Wilk normality test on the following variables: `Ozone`, `Temp`, `Solar.R`, and `Wind.`

    ```{r}
      shapiro.test(airquality$Ozone) 
    #data:  airquality$Ozone
    #W = 0.87867, p-value = 2.79e-08
      shapiro.test(airquality$Temp)
      #data:  airquality$Temp
    #W = 0.97617, p-value = 0.009319
      shapiro.test(airquality$Solar.R)
      #data:  airquality$Solar.R
    #W = 0.94183, p-value = 9.492e-06
      shapiro.test(airquality$Wind)
      #data:  airquality$Wind
    #W = 0.98575, p-value = 0.1178
    ```

3.  What is the purpose of the Shapiro-Wilk test?

    1.  The purpose of the Shapiro-Wilk test is to determine whether a data set is normally distributed or not. This statistical analysis works best on small data sets as it is highly sensitive to deviations from a mean.

4.  What are the null and alternative hypotheses for this test?

    1.  Null: Data is normally distributed

    2.  Alternative: The data is not normally distributed

5.  Interpret the p-values. Are these variables normally distributed?

    1.  The temperature analysis has a p-value of 0.009 which is smaller than 0.05 and is statistically significant because it assumes there is a very low chance of observing data as extreme under the conditions of the null. The same conclusion can be drawn from the p-values of ozone and solar radiation as the values of these tests are much smaller than 0.05. The p-value of the wind test is higher than 0.05 which indicates there is a high probability of the same results occurring under the conditions of the null. This value is considered statistically insignificant.

        ```         
        data:  airquality$Ozone
        W = 0.87867, p-value = 2.79e-08

        data:  airquality$Temp
        W = 0.97617, p-value = 0.009319

        data:  airquality$Solar.R
        W = 0.94183, p-value = 9.492e-06

        data:  airquality$Wind
        W = 0.98575, p-value = 0.1178
        ```

## **Part 2: Data Transformation and Feature Engineering**

6.  Create a new column with `case_when` translating the Months into four seasons (Winter (Nov, Dec, Jan), Spring (Feb, Mar, Apr), Summer (May, Jun, Jul), and Fall (Aug, Sep, Oct)).

    ```{r}
install.packages("dplyr") # Install if needed
library(dplyr)
seasonality <- airquality %>%
      mutate(season = case_when(
        Month %in% c(11, 12, 1) ~ "Winter",
        Month %in% c(2, 3, 4) ~ "Spring",
        Month %in% c(5, 6, 7) ~ "Summer",
        Month %in% c(8, 9, 10) ~ "Fall"))

    head(seasonality)
    ```

7.  Use `table` to figure out how many observations we have from each season.

    ```{r}
    table(seasonality$season)
    ```

## **Part 3: Data Preprocessing**

8.  Normalize the predictor variables (Temp, Solar.R, Wind, and Season) using a `recipe`
```{r}
recipes::step_normalize(airquality$Temp)
```


9.  What is the purpose of normalizing data?
  1.

10. What function can be used to impute missing values with the mean?
  1.

11. `prep` and `bake` the data to generate a processed dataset.
```{r}

```


12. Why is it necessary to both `prep()` and `bake()` the recipe?
  1. 

## **Part 4: Building a Linear Regression Model**

13. Fit a linear model using Ozone as the response variable and all other variables as predictors. Remeber that the `.` notation can we used to include all variables.
```{r}

```


14. Interpret the model summary output (coefficients, R-squared, p-values) in plain language
  1.

## **Part 5: Model Diagnostics**

15. Use `broom::augment` to suppliment the normalized `data.frame` with the fitted values and residuals.
```{r}

```


16. Extract the residuals and visualize their distribution as a histogram and qqplot.
```{r}

```


17. Use `ggarange` to plot this as one image and interpret what you see in them.
```{r}

```


18. Create a scatter plot of actual vs. predicted values using ggpubr with the following setting:
```{r}

```

```         
ggscatter(a, x = "Ozone", y = ".fitted",
          add = "reg.line", conf.int = TRUE,
          cor.coef = TRUE, cor.method = "spearman",
          ellipse = TRUE)
```

19. How strong of a model do you think this is?
  1.

20. Render your document to HTML and submit to Canvas. Be sure to include this in your document yml:
