---
title: "exercise-22-csu"
author: "Eleanor Lindsey"
format:
  html:
    self-contained: true
execute:
  echo: true
editor: visual
---

#### Load in Necessary Libraries and Data

```{r}
# Load libraries
options(repos = c(CRAN = "https://cran.rstudio.com/"))
library(dataRetrieval)
library(dplyr)
library(tsibble)
library(tidymodels)
library(modeltime)
library(lubridate)
library(timetk)
library(prophet)
library(forecast)
library(zoo)

# Download daily streamflow and aggregate to monthly
poudre_flow <- readNWISdv(siteNumber = "06752260",
                          parameterCd = "00060",
                          startDate = "2013-01-01",
                          endDate = "2023-12-31") %>%
  renameNWISColumns() %>%
  mutate(Date = as.Date(Date)) %>%
  mutate(month = floor_date(Date, "month")) %>%
  group_by(month) %>%
  summarise(Flow = mean(Flow, na.rm = TRUE), .groups = "drop") %>%
  rename(Date = month)

# Convert to tsibble
poudre_tsibble <- as_tsibble(poudre_flow, index = Date)


```

#### Model Making

```{r}
# Split the data
splits <- initial_time_split(poudre_tsibble, prop = 0.92)

# Define model specs
poudre_arima_model <- arima_reg() %>% set_engine("auto_arima")
poudre_prophet_model <- prophet_reg() %>% set_engine("prophet")

# Fit models
model_tbl <- modeltime_table(
  fit(poudre_arima_model, Flow ~ Date, data = training(splits)),
  fit(poudre_prophet_model, Flow ~ Date, data = training(splits))
)

```

#### Get 2024 Daily Streamflow and Aggregate to Monthly

```{r}
# Download 2024 observed data
poudre_2024 <- readNWISdv(siteNumber = "06752260",
                          parameterCd = "00060",
                          startDate = "2024-01-01",
                          endDate = "2024-12-31") %>%
  renameNWISColumns() %>%
  mutate(Date = as.Date(Date)) %>%
  mutate(month = floor_date(Date, "month")) %>%
  group_by(month) %>%
  summarise(observed_cfs = mean(Flow, na.rm = TRUE), .groups = "drop") %>%
  rename(Date = month)

```

#### Forecast Next 12 Months and Aggregate to Monthly

```{r}
# Refit models on full data
refit_tbl <- model_tbl %>% modeltime_refit(poudre_tsibble)

# Create 12 months of future dates
last_date <- max(poudre_tsibble$Date)
future_dates <- tibble(Date = seq.Date(from = last_date %m+% months(1),
                                       by = "month", length.out = 12))

# Generate forecasts
forecast_tbl <- refit_tbl %>%
  modeltime_forecast(new_data = future_dates, actual_data = poudre_tsibble)

# Aggregate forecasts to monthly means
forecast_monthly <- forecast_tbl %>%
  filter(.key == "prediction") %>%
  mutate(month = floor_date(.index, "month")) %>%
  group_by(month) %>%
  summarise(predicted_cfs = mean(.value, na.rm = TRUE), .groups = "drop") %>%
  rename(Date = month)

```

#### Compare Predictions with Observed Data & Calculate R²

```{r}
# Join observed and predicted data
comparison_tbl <- left_join(forecast_monthly, poudre_2024, by = "Date") %>%
  filter(!is.na(observed_cfs), !is.na(predicted_cfs))

# Compute R² using linear model
if (nrow(comparison_tbl) > 1) {
  r2_model <- lm(observed_cfs ~ predicted_cfs, data = comparison_tbl)
  r2_value <- summary(r2_model)$r.squared
  cat("R² between predictions and observed values:", round(r2_value, 3), "\n")
} else {
  cat("Not enough overlapping data to compute R² — check monthly alignment.\n")
}

```

R\^2 value of 0.85 which suggests the model is credible and produces reliable predictions.

#### Last, generate a plot of the Predicted vs Observed values and include a 1:1 line, and a linear model line.

```{r}
# Plot predicted vs. observed
poudre_comparison_plot<-ggplot(comparison_tbl, aes(x = predicted_cfs, y = observed_cfs)) +
  geom_point(color = "darkgreen", size = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
  geom_smooth(method = "lm", se = FALSE, color = "blue", linewidth = 1) +
  labs(
    title = "Predicted vs Observed Streamflow (Monthly Averages, 2024)",
    x = "Predicted Flow (cfs)",
    y = "Observed Flow (cfs)"
  ) +
  theme_minimal(base_size = 14)

poudre_comparison_plot

```
